Evolution of the AGN UV luminosity function

### 2017-07-06 Paper/paper.tex
Add photoionisation rate plot for two luminosity cutoffs
Update figures
Update text

\documentclass[a4paper,fleqn,usenatbib]{mnras}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	
\usepackage{amsmath}	
\usepackage{amssymb}
\usepackage{pdflscape}
\usepackage{siunitx}

@@

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.}
  \label{fig:gammapi}
\end{figure*}

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.6\textwidth,keepaspectratio]{gammapi.pdf}
  \end{center}
  \caption{Hydrogen photoionization rate}
  \label{fig:gammapi}
\end{figure*}

>>>>

\documentclass[a4paper,fleqn,usenatbib]{mnras}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	
\usepackage{amsmath}	
\usepackage{amssymb}
\usepackage{pdflscape}
\usepackage{siunitx}
\def\lya{Ly$\alpha$~}

@@

\section{Introduction}

The luminosity function of active galactic nuclei (AGN) and its
evolution over cosmological time scales has been a matter of central
interest of a large body of work over the last four decades
\citep[e.g.,][]{1978A&A....68...17M, 1983ApJ...269..352S,
  1988ApJ...325...92K, 1988MNRAS.235..935B, 1993ApJ...406L..43H,
  1994ApJ...421..412W, 1995AJ....110...68S, 1995AJ....110.2553K,
  1995ApJ...438..623P, 2000MNRAS.317.1014B, 2001AJ....121...54F,
  2004AJ....128..515F, 2006AJ....131.2766R, 2007ApJ...654..731H,
  2009MNRAS.392...19C, 2010AJ....139..906W, 2011ApJ...728L..26G,
  2013ApJ...773...14R, 2013ApJ...768..105M, 2015AA...578A..83G,
  2015ApJ...798...28K, 2016ApJ...829...33Y, 2016ApJ...833..222J,
  2017MNRAS.466.1160M}.  Determination of the AGN luminosity function
constrains models of the build-up of supermassive black holes
\citep{2015MNRAS.452..575S, 2016MNRAS.462..190R}.  Due to the
incidence of supermassive black holes in most galaxies, the tight
scaling relations observed to exist between the mass of these black
holes and properties of their host galaxies
\citep{2013ARA&A..51..511K, 2013ApJ...764..184M}, and the increasing
consensus that AGN activity feeds back on the host galaxy evolution,
the AGN luminosity function also constrains models of galaxy
formation.  Finally, thanks to their relative brightness and high
Lyman continuum escape fraction, the luminosity function of AGN
determines their influence on the temperature and ionization state of
the intergalactic medium (IGM) over large scales, possibly even
primarily driving hydrogen and helium reionization.

The claimed discovery of 19 low-luminosity ($M_{1450}>-22.6$) AGN
between redshifts $z=4.1$ and $6.3$ by \citet{2015AA...578A..83G}
using a novel X-ray/NIR selection criterion, has renewed the interest
in the evolution of the UV luminosity function of AGN.  This finding
suggested that the faint end of the quasar UV luminosity function is
steeper at these redshifts than previously thought
\citep{2007ApJ...654..731H, 2012ApJ...746..125H}.  Using far-UV
spectral slopes from composite spectra of low-redshift quasars, and
assuming a Lyman continuum escape fraction of 100\%,
\citet{2015AA...578A..83G} argued that AGN brighter than
$M_{1450}=-18$ can potentially produce all of the metagalactic
hydrogen photoionization rate inferred from the \lya forest at
$4<z<6$.  Additional indications of a significant presence of AGN at
high redshift ($z\sim 6$) have emerged.  First,
\citet{2015MNRAS.447.3402B} reported a large scatter in the \lya
opacity between different sightlines close to redshift
$z=6$. \cite{2015MNRAS.453.2943C} showed that these opacity
fluctuations extend to substantially larger scales ($\gtrsim 50\,
h^{-1}$cMpc) than expected in reionization histories dominated by
low-luminosity galaxies (see also \citealt{2016MNRAS.460.1328D}).
\citet{2017MNRAS.465.3429C} further demonstrated that opacity
fluctuations on such large scales arise naturally if there is a
significant contribution ($\gtrsim 50\%$) of AGN to the ionising
emissivity at the redshift of the observed opacity fluctuations ($z
\sim 5.5$--$6$) as would be expected for a QSO luminosity that is
consistent with the measurements of \citet{2015AA...578A..83G}.
Second, measurements of the Lyman continuum escape fraction from
high-redshift galaxies are still elusive.  Although high-redshift
galaxies as faint as rest-frame UV magnitude $M_\mathrm{UV}=-12.5$
($L\sim 10^{-3}L^*$) at $z=6$ \citep{2017ApJ...835..113L} and
redshifts as high as $z=11.1$ \citep{2016ApJ...819..129O} have now
been reported, the escape of Lyman continuum photons has been detected
in only a small number of comparatively bright ($L>0.5L^*$)
low-redshift ($z < 4$) galaxies.  In these galaxies, the escape
fraction is typically found to be 2--20\% \citep{2010ApJ...725.1011V,
  2011ApJ...736...41B, 2015ApJ...804...17S, 2015ApJ...810..107M,
  2016A&A...585A..48G, 2017MNRAS.468..389J, 2017MNRAS.465..316M} but
reionization would require escape fraction of about 20\% in galaxies
down to $M_\mathrm{UV}=-13$ \citep{2016PASA...33...37F,
  2015ApJ...802L..19R, 2016MNRAS.457.4051K}.  Finally, incidence of
high-redshift AGN is also consistent with the shallow bright-end
slopes of the high-redshift ($z\sim 7$) UV luminosity function of
galaxies relative to a Schechter-function representation
\citep{2012MNRAS.426.2772B, 2014MNRAS.440.2810B, 2014ApJ...792...76B,
  2015MNRAS.452.1817B} and the hard spectra of these bright galaxies
\citep{2015MNRAS.450.1846S, 2015MNRAS.454.1393S, 2017MNRAS.464..469S}.

However, these arguements in favour of a higher incidence of
UV-emitting AGN at high redshit, have been contested by several
observations.  \citet{2016arXiv160706467D} considered the effect of
AGN-dominated reionization on the Ly$\alpha$ opacity at $z>5$,
He~\textsc{ii} Ly$\alpha$ opacity at $z\sim 3.1$--$3.3$, and the
thermal history of the IGM.  In agreement with
\citet{2015MNRAS.453.2943C}, these authors found that AGN did provide
a plausible explanation for the large fluctuations in the Ly$\alpha$
opacity at $z>5$.  However, they found that reionization of
He~\textsc{ii} occurs much earlier in these AGN-dominated models (see
also \citealt{2016arXiv160602719M}).  For instance, in the model of
\citet{2015ApJ...813L...8M}, He~\textsc{ii} reionization is complete
at $z=4.5$, compared to $z=3$ in the standard scenario
\citep{2012ApJ...746..125H}.  This early Helium reionization could
result in higher IGM temperatures due to the associated photoheating.
The temperature of the IGM at mean density is enhanced in
AGN-dominated models by factors of $\sim 2$ relative to the standard
models for $z=3.5$--$5$, in conflict with measurements.  This
inconsistency could be avoided by reducing the escape fraction of 4~Ry
photons in AGN, but it is not clear if this can be achieved while
requiring a 100\% escape fraction of 1~Ry photons in order to explain
the Ly$\alpha$ opacity fluctuations.  Further evidence against
AGN-dominated reionization models has emerged from metal-line
absorbers at $z\sim 6$.  In their cosmological radiation
hydrodynamical simulations, \citet{2016MNRAS.459.2299F} find that the
hard spectral slopes of UV backgrounds in AGN-only reionization models
produce too many C~\textsc{iv} absorption systems relative to
Si~\textsc{iv} and C~\textsc{ii} at $z\sim 6$.  However, these
simulations assume an $L_\nu\propto\nu^{-1.57}$ AGN SED at extreme UV
\citep{2001AJ....122..549V, 2002ApJ...579..500T, 2012ApJ...746..125H}.
This slope is marginally harder than recent measurements
($L_\nu\propto\nu^{-1.7}$) from a stack of $z\sim 2.4$ quasars
\citep{2015MNRAS.449.4204L}.  \citet{2016MNRAS.459.2299F} also find
that the N(Si~\textsc{iv})/N(C~\textsc{iv}) column density ratio
measurements prefer a somewhat harder and more intense $>4$~Ry
background than the standard model of \citet{2012ApJ...746..125H}.
Using a large sample of X-ray-selected quasars in the redshift range
$z=0$--$6$, \citet{2017MNRAS.465.1915R} find that the faint end of the
AGN UV luminosity function at $z\sim 6$ is likely to be much shallower
that that reported by \citet{2015AA...578A..83G}.  In their analysis,
\citet{2017MNRAS.465.1915R} use an AGN obscuration optical depth
($\log N_\mathrm{H}$) cut-off that reproduces low-redshift AGN UV
luminosity functions and an X-ray-to-optical/UV luminosity ratio
calibrated at redshifts $z=0.05$--$4$ \citep{2010A&A...512A..34L}.
These authors argue that the apparent contradiction with the results
of \citet{2015AA...578A..83G} could be explained by contamination
from the AGN host galaxies.  It has also been recently argued that the
Lyman continuum escape fraction of AGN might not be 100\% as is
usually assumed \citep{2017MNRAS.465..302M}.  This may further reduce
the contribution of AGN to reionization.

We assume a flat cosmology with density parameters
$\left(\Omega_\mathrm{m},\Omega_\Lambda\right)=\left(0.3,0.7\right)$
and a Hubble constant $H_0=70$\,km\,s$^{-1}$\,Mpc$^{-1}$. Magnitudes
are reported in the AB system \citep{1983ApJ...266..713O}, and
observed magnitudes are point spread function (PSF) magnitudes
\citep{2002AJ....123..485S} corrected for Galactic extinction
\citep{1998ApJ...500..525S} unless otherwise noted.

\section{Homogenised Quasar Sample}

\subsection{Sample Selection}

We started with compiling the samples of recent UV-optical quasar
surveys. The restriction to UV-optical surveys was mainly driven by
our science goal to characterise the UV luminosity function of Type~I
quasars, but also by the still uncertain conversion from X-ray flux to
UV flux, which adds unnecessary systematics for large
samples. \textbf{References!} The individual surveys and their main
characteristics are listed in
Table~\ref{tab:samples}. Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item Spectroscopic redshifts for the majority of targets.
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or $M_{1450}$).
\item A characterised selection function.
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or from the authors. As a reference for future surveys we
publish them here in modified and homogenised form
(Section~\ref{sect:datahom}). \textbf{Maybe get the authors' consent.}

Due to their selection criteria and their statistical power specific
surveys contribute to distinct redshift ranges. At $z<2.2$ we
considered quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and the 2SLAQ catalogue
\citep{2009MNRAS.392...19C}. We restricted the SDSS DR7 sample to XXX
$z<2.2$ quasars selected with the final SDSS quasar selection
algorithm \citep{2002AJ....123.2945R, 2006AJ....131.2766R} from a
survey area of 6248\,deg$^2$ \citep{2012ApJ...746..169S}. We adopted
the SDSS targeting photometry corrected for Galactic extinction
\citep{2010AJ....139.2360S}. To limit systematic uncertainties in the
correction for host galaxy light \citep{2009MNRAS.392...19C} we
restricted the 2SLAQ sample to XXX $g<21.85$ $0.4<z<2.2$ quasars from
its spectroscopic survey footprint near the North Galactic Pole (NGP,
XX quasars in $127.7$\,deg$^2$) and the South Galactic Pole (SGP, XX
quasars in $64.2$\,deg$^2$). The small sample overlap between SDSS and
2SLAQ (112 quasars) has negligible impact on the QLF evaluation.

At $2.2<z<3.5$ we used a single sample of 23,301 uniformly
colour-selected quasars from 2236\,deg$^2$ in BOSS DR9
\citep{2013ApJ...773...14R} due to several improvements compared to
previous surveys. First, it covers a similar magnitude range as 2SLAQ
but with $>20$ times as many quasars. Second, although the SDSS DR7
sample provides better coverage of the bright end of the QLF at these
redshifts, its selection function is highly dependent on the assumed
incidence of (partial) Lyman limit systems in the IGM
\citep{2009ApJ...705L.113P, 2011ApJ...728...23W}. While the BOSS DR9
selection function considers these improvements, the uncertainty in
the QLF remains dominated by assumptions in the selection function
given the large sample size
\citep{2013ApJ...773...14R}. Variability-selected quasar samples
circumvent this issue \citep{2013ApJ...773...14R, 2013A&A...551A..29P,
  2016A&A...587A..41P}, but may be affected by (i) single-epoch
imaging incompleteness at the faint end \citep{2013ApJ...773...14R},
and (ii) uncertainties in the selection function caused by the limited
number of known $z\ga 3$ quasars not selected by variability in the
same footprint \citep{2013A&A...551A..29P, 2016A&A...587A..41P}.

At $3.7<z<4.1$ we used a combination of SDSS DR7 \citep[XX uniformly
  selected quasars from][]{2010AJ....139.2360S} and the NDWFS$+$DLS
survey \citep{2010ApJ...710.1498G,2011ApJ...728L..26G}. The lower cut
$z>3.7$ limits the impact of systematic uncertainties in the
\citet{2006AJ....131.2766R} SDSS selection function
\citep{2009ApJ...705L.113P, 2011ApJ...728...23W}. We did not consider
the results from surveys for faint $z\sim 4$ quasars in the COSMOS
field \citep{2011ApJ...728L..25I, 2012ApJ...755..169M} due to
systematic errors in their selection functions\footnote{Both studies
  simulated quasar colours with a mean IGM attenuation curve
  \citep{1995ApJ...441...18M} that cannot account for stochastic Lyman
  continuum absorption, and therefore underpredicts the variance in
  quasar colours \citep{1999ApJ...518..103B, 2008MNRAS.387.1681I,
    2011ApJ...728...23W}. Modelling the colour variance in these
  surveys is essential, as most of the \citet{2011ApJ...728L..25I}
  quasars are near the edge of their colour selection region (see
  their Figure~1), and \citet{2012ApJ...755..169M} require modest
  attenuation of the $U$ band flux relative to the mid-infrared
  flux.}. Furthermore, 30 per cent of the \citet{2012ApJ...755..169M}
COSMOS sample have visually estimated photometric redshifts, and the
spectroscopic subsample reveals that 40 per cent of the visually
estimated redshifts are biased low
($z_\mathrm{spec}>z_\mathrm{est}+0.3$, see their Figure~9). These
unaccounted systematic redshift errors at least partly explain the
discrepancy in the $z\sim 4$ QLF between \citet{2011ApJ...728L..26G}
and \citet{2012ApJ...755..169M}, which justifies our preference for
the former sample.

Between $z=4.1$ and $z=4.7$ our combined sample has XX uniformly
selected quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and XX quasars from
\citet{2011ApJ...728L..26G}. Although we do not include the highly
debated \citet{2015AA...578A..83G} sample in our main analysis due to
its rough selection function and lacking spectroscopy for 17 of the 22
quasar candidates, we use it in Section~XX to constrain the faint end
($M_{1450}>-23$) of the QLF at $z>4.1$. We restricted the
\citet{2015AA...578A..83G} sample to the 19/22 sources considered in
their QLF.

At $4.7\le z<5.1$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 71 $z<5.1$ quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} together with its selection function.
%Although \citet{yang} are sensitive to $z>5.1$ quasars, we limit
%their sample to $z<5.1$ to limit differential evolution of the QLF
%across the redshift range.
To avoid double-counting quasars contained both in
\citet{2016ApJ...829...33Y} and the SDSS DR7 sample from
\citet{2013ApJ...768..105M}, we used the latter only at
$M_{1450}>-26.72$ (in our adopted cosmology), yielding 94 additional
$4.7\le z<5.1$ quasars selected in 6248\,deg$^2$. We used the updated
$z\sim 5$ SDSS DR7 selection function \citep{2013ApJ...768..105M}
instead of the one from \citet{2006AJ....131.2766R}.
%\footnote{We note that the major difference in the selection
%functions must be due to different input parameters, since the single
%color selection criterion in \citet{mcgreer} selects the vast
%majority of the SDSS DR7 quasars.}.
To these two bright-end samples we added the faint-end sample from the
\citet{2013ApJ...768..105M} SDSS Stripe~82 survey (52 uniformly
selected $4.7\le z<5.1$ quasars in 235\,deg$^2$) and two quasars from
\citet{2011ApJ...728L..26G}, adopting the respective selection
functions. We did not consider the limit on the $z\sim 5$ QLF by
\citet{2012ApJ...756..160I} due to systematic errors in their
selection function\footnote{\citet{2012ApJ...756..160I} underestimated
  the dispersion in rest-frame UV quasar colours with respect to SDSS
  at all redshifts (their Figure~4). Contrary to their claim,
  photometric errors have a small effect on the colour distribution of
  SDSS quasars given the statistical errors of $<0.03$\,mag in $gri$
  for 90 per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and
  a relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

The SDSS colours of $5.1\le z<5.5$ quasars are similar to those of M
and L dwarf stars, resulting in a low and uncertain selection
efficiency \citep{2013ApJ...768..105M}. WISE mid-infrared selection
performs better \citep{2016ApJ...829...33Y}, but is restricted to the
bright end of the quasar population. Our combined sample contains the
remaining 28 quasars from \citet{2016ApJ...829...33Y} and 10 quasars
from the \citet{2013ApJ...768..105M} Stripe~82 survey. As we will show
in Section~XX, the resulting QLF is consistent with those at lower and
higher redshifts, indicating that the \citet{2013ApJ...768..105M}
selection functions are quite reliable. \textbf{Check this!}

Finally, we combined the samples from all spectroscopic $z\sim 6$
quasar surveys with a determined selection function as of June
2017. \citet{2016ApJ...833..222J} recently compiled all quasars
discovered in several SDSS $z\sim 6$ surveys together with
consistently derived selection functions. Their uniform sample
consists of 24 quasars from the SDSS main survey (11240\,deg$^2$), 10
additional quasars in regions with two or more SDSS imaging scans
(so-called overlap regions, 4223\,deg$^2$), and 13 faint quasars from
SDSS Stripe~82 (277\,deg$^2$). The CFHQS \citep{2010AJ....139..906W}
provided a uniform sample of 16 quasars in the Very Wide Survey
(494\,deg$^2$) and a single quasar in the Deep Survey
($4.47$\,deg$^2$). The one quasar detected in both SDSS and CFHQS does
not lead to underestimated statistical errors in the QLF. Lastly, we
included the two objects from \citet{2015ApJ...798...28K}. One might
be a Lyman break galaxy due to its narrow Ly$\alpha$ emission line
(half width at half maximum 427\,km\,s$^{-1}$), but
\citet{2015ApJ...798...28K} note that 10/17 candidates do not have
follow-up spectroscopy, hence their QLF may be underestimated. Since
very small samples preclude a statistical correction via the selection
function, complete spectroscopic follow-up of $z\sim 6$ candidates is
essential. Although we will account for the slightly different
redshift sensitivities for the different surveys, we will quote a
nominal redshift range $5.7<z<6.5$ for the combined $z\sim 6$ sample.

%\clearpage
\subsection{Sample Homogenisation}
\label{sect:datahom}

For a joint fit of the QLF it is necessary to homogenise the different
survey samples in absolute magnitude, and to convert their selection
functions to the same absolute magnitude system. For the analysis of
the quasar UV emissivity and to be consistent with published work at
$z>3$ we chose to convert all samples and selection functions to the
absolute AB magnitude at a rest frame wavelength $\lambda=1450$\,\AA
\begin{equation}\label{eq:absmag}
M_{1450}\left(z\right) = m-5\log{\left(\frac{d_L\left(z\right)}{\mathrm{Mpc}}\right)}-25-K_{m,1450}\left(z\right),
\end{equation}
with the luminosity distance $d_L\left(z\right)$ to a quasar at
redshift $z$, the apparent magnitude $m$ in a filter used in the
survey, and the bandpass correction $K_{m,1450}\left(z\right)$
\citep{1956AJ.....61...97H, 1968ApJ...154...21O, 2000A&A...353..861W,
  2002astro.ph.10394H}. For the bandpass correction we used a
combination of the \citet{2015MNRAS.449.4204L} quasar SED at
$\lambda<2500$\,\AA, and the \citet{2001AJ....122..549V} quasar
composite spectrum at longer wavelengths to cover the lowest
redshifts.  The samples from SDSS and BOSS are defined in the SDSS $i$
band, while 2SLAQ is defined in the $g$ band. At $z>4.7$ we adopted
the SDSS $z$ band magnitude (in the following denoted $z_\mathrm{AB}$)
for SDSS DR7 quasars to avoid additional corrections due to the
Ly$\alpha$ forest. Figure~XX shows our bandpass corrections for SDSS,
BOSS and 2SLAQ as a function of redshift. While the
\citet{2015MNRAS.449.4204L} SED is for bright ($M_{1450}\simeq -27.2$)
quasars, UV composite spectra including fainter quasars
\citep{2002ApJ...565..773T, 2012ApJ...752..162S, 2014ApJ...794...75S}
give similar values, such that our bandpass corrections likely remain
applicable at $M_{1450}\la -24$. \textbf{What about even fainter AGN?}

For the $z<2.2$ sample we corrected the SDSS $i$ and 2SLAQ $g$ band
magnitudes for host galaxy contamination following
\citet{2009MNRAS.392...19C}. Considering the different magnitude
limits of 2SLAQ and SDSS, the modelled host galaxy contamination is
small for $z>0.5$ quasars ($<0.1$\,mag in $g$, $<0.2$\,mag in $i$),
and is negligible at $z>0.8$.  In case the band defining the magnitude
limit of the survey undesirably overlaps with the Ly$\alpha$ forest
\citep{2010ApJ...710.1498G, 2011ApJ...728L..26G, 2013ApJ...768..105M}
we adopted their respective bandpass corrections to $M_{1450}$. In
particular, for the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G}
sample we recomputed $M_{1450}$ from the $R$ band photometry to be
consistent with the selection function defined in $R$, and to avoid
uncertainties in their spectrophotometry due to incomplete spectral
coverage. For the remaining high-redshift surveys reporting $M_{1450}$
obtained by various methods \citep{2010AJ....139..906W,
  2015ApJ...798...28K, 2015AA...578A..83G, 2016ApJ...829...33Y,
  2016ApJ...833..222J} we did not re-compute $M_{1450}$, but applied
appropriate shifts to correct to our adopted cosmology.

The selection functions were treated similarly, i.e.\ the photometric
selection function of survey $j$ given in observed magnitudes
$f_{\mathrm{p},j}\left(m,z\right)$ \citep{2006AJ....131.2766R,
  2009MNRAS.392...19C, 2010ApJ...710.1498G, 2013ApJ...773...14R} was
transformed to our absolute magnitudes
$f_{\mathrm{p},j}\left(M_{1450},z\right)$ with
Equation~\ref{eq:absmag}, while the ones given in $M_{1450}$ were
adjusted to our cosmology. Note, however, that many surveys report
additional sources of incompleteness that preclude a straightforward
re-evaluation of the QLF.

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$
\citep{2013ApJ...773...14R}, and we adopted
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$. Finally,
we obtained a rough estimate of the \citet{2015AA...578A..83G}
selection function by taking
$f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$ (see their
Table~3).

% BOSS:
% - color selection function in Table 5 of Ross et al. (imag, z), sample contains QSOs with spectrocoverage completeness >0.85, mean spectrocoverage completeness 0.962 multiplied onto color selection function
% - interpolate Beta's K correction (imag->1450A) and compute M1450 for our cosmology (sample and map)

% 2SLAQ:
% - low-luminosity AGN -> model and subtract host galaxy contribution, magnitudes only for nuclear component, host galaxy contribution <20% at z>0.4 for faintest sources in 2SLAQ (Croom et al. 2009, MNRAS, 392, 19)
% - several selection functions (morphological, color, spectroscopic coverage, spectroscopic success)
% - Table 12: color selection function (available online), gmag is total PSF mag including host galaxy contribution, subtract gmag host galaxy contribution listed in Table 12, selection function extrapolated at bright end (lacking coverage of whole catalog)
% - coverage completeness for NGP and SGP given in Fig 4 as a function of gmag, interpolated and multiplied into color selection function
% - spectroscopic success as function of gmag in Fig 6b (only quality=1 redshifts considered), interpolated and multiplied onto color selection function
% - convert g_agn to M1450 with L15 SED (valid for bright quasars, check with K correction for Telfer/Shull)
% - sample: correct PSF gmag for Galactic extinction, subtract host galaxy light (Table 12), interpolate Beta's K correction (gmag->1450A) and compute M1450 for our cosmology
% SDSS:
% - Richards et al. selection function given in (z,imag) Fig 6 and Table 1, corrected for edge effects at magnitude limits (steep gradient) and NaN values, selection function corrected for imaging incompleteness (factor 0.95)
% - imag in selection function converted to M1450 with L15 SED, conversion uncertain at z>4.7 due to Lya forest
% - sample: targeting psf imag corrected for Galactic extinction (Schlegel) and converted to M1450 with L15 SED
% - z<2.2: sample limited to i_dered<19.1, imag of sample and selection function corrected for host galaxy light (Table 12 in Croom et al 2009, given for imag and gmag)
% Glikman:
% - different selection functions for 2 sub-surveys, selection functions corrected for interpolation errors (values >1 reset to 1), selection functions given in (z,Rmag) Fig 7 of G10, converted to M1450 with their Rmag K correction (Fig 8 in G10)
% - spectroscopic incompleteness (f_spec, given in G10 and updated in G11) incorporated into selection functions via linear fit to f_spec(Rmag) at Rmag>23 and constant otherwise
% - sample M1450 recomputed from Rmag and K correction to be consistent with completeness map, M1450=Rmag-Kcorr-5*log10(d_L/Mpc)-25
% McGreer:
% - Stripe 82: selection function (Fig. 11) modified to include photometric incompleteness (factor 0.95) and spectroscopic incompleteness (imag>20.8, Fig. 14), M1450 in selection map and sample M1450 converted to our cosmology (adding 0.07)
% - DR7: sample not listed in paper, M1450 computed from zmag (QSO continuum) and L15 SED, same procedure applied to S82 reveals L15 magnitudes are slightly brighter than M13 magnitudes (median 0.12mag) -> small change,
%        M13 selection function corrected for photometric (factor 0.95) and spectroscopic (factor 0.95) incompleteness
% Yang: selection function (Fig. 5) modified to include image selection incompleteness (factor 0.97) and spectroscopic incompleteness (factor ~0.84 at zmag>19, Fig 7), M1450 in selection map and sample M1450 converted to our cosmology
% Willott: selection map M1450 adjusted to our cosmology adding 0.05, slightly different cosmological parameters in Willott et al. (2007,2009,2010) -> M1450 converted
% Jiang: selection functions unchanged, M1450 as given in paper
% Kashikawa: map not corrected for spectroscopic incompleteness (sample size), M1450 as given in paper
% Giallongo: selection function estimated from phi_obs and phi_corr (Table 3 of Giallongo et al.), M1450 as given in paper

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.}
  \label{fig:emissivity}
\end{figure*}

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{g_18.pdf}
    \includegraphics[width=0.5\textwidth]{g_20.pdf}
    \end{tabular}
  \end{center}
  \caption{AGN contribution to the hydrogen photoionisation rate,
    assuming 100\% escape fraction.  Model luminosity functions are
    integrated down to $M_{1450}=-18$ in the left panel and $-20$ in
    the right panel.}
  \label{fig:gammapi}
\end{figure*}
